---
title: "Development of Renewable Energy Share in Germany Compared to Global Trends (1990–2020)"
author: "Ayoub Jebali"
output: 
  pdf_document:
    toc: true
    toc_depth: 2
    number_sections: true
---

# Introduction
This analysis examines the development of the renewable energy share in total energy consumption in Germany compared to the global average from 1990 to 2020. The focus is on whether Germany takes a pioneering role or whether other countries/regions have made similar or even faster progress.

# Datasets
The datasets come from the World Bank Group:
"Renewable energy consumption (% of total final energy consumption)"

This dataset provides comprehensive data on renewable energy consumption as a percentage of total final energy consumption across different countries and regions from 1990 to 2020.

\newpage

# Code Section
## Reading Data

```{r setup, echo=TRUE, message=FALSE, warning=FALSE}
# Load libraries
library(dplyr)
library(ggplot2)
library(readr)
library(reshape2)
library(mosaic)

# Read data
energy_data <- read_csv("Renewable_energy_consumption_worldwide.csv",
                        col_types = cols(`Indicator Name` = col_skip(),
                                         `Indicator Code` = col_skip(),
                                         `Country Code` = col_skip(),
                                         ...69 = col_skip()))

# Overview of the data
head(energy_data)
# Using inspect(energy_data) or other functions would be too long
```
\newpage

## Data Cleaning

```{r}

# Check for missing data
sum(is.na(energy_data))

# Identify numeric columns corresponding to years 1960-1989
columns_to_remove <- which(colnames(energy_data) %in% as.character(1960:1989))
# Remove these columns
energy_data <- energy_data[, -columns_to_remove]

# Identify columns corresponding to years 2021 and later
columns_to_remove <- which(colnames(energy_data) %in% as.character(2021:2100))
# Remove these columns
energy_data <- energy_data[, -columns_to_remove]

# Check for missing data
sum(is.na(energy_data))

# Overview of the data
head(energy_data)
```
\newpage
```{r}
# Transform data to long format
energy_data_long <- melt(energy_data,
                         id.vars = c("Country Name"),
                         variable.name = "Year",
                         value.name = "Renewable_Energy_Share")

# Convert year to numeric value
energy_data_long$Year <- as.numeric(as.character(energy_data_long$Year))

# Renewable energy share in Germany
germany_data <- energy_data_long %>%
  filter(`Country Name` == "Germany")

# Overview of data for Germany
head(germany_data)
favstats(~Renewable_Energy_Share, data=germany_data)

# Filter global average
worldwide_avg <- energy_data_long %>%
  filter(`Country Name` == "World")

# Overview of global average values
head(worldwide_avg)
favstats(~Renewable_Energy_Share, data=worldwide_avg)
```
\newpage

## Data Visualization

```{r, fig.width=10, fig.height=6}
# Create main comparison plot
ggplot() +
  geom_line(data = germany_data, 
            aes(x = Year, y = Renewable_Energy_Share, color = "Germany"), 
            size = 1.2) +
  geom_line(data = worldwide_avg, 
            aes(x = Year, y = Renewable_Energy_Share, color = "Global"), 
            size = 1.2) +
  geom_point(data = germany_data, 
             aes(x = Year, y = Renewable_Energy_Share, color = "Germany"), 
             size = 0.8, alpha = 0.7) +
  geom_point(data = worldwide_avg, 
             aes(x = Year, y = Renewable_Energy_Share, color = "Global"), 
             size = 0.8, alpha = 0.7) +
  labs(title = "Development of Renewable Energy Share (1990-2020)",
       subtitle = "Germany vs. Global Average",
       x = "Year",
       y = "Renewable Energy Share (%)",
       color = "Region") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5, size = 14, face = "bold"),
        plot.subtitle = element_text(hjust = 0.5, size = 12),
        legend.position = "bottom") +
  scale_color_manual(values = c("Germany" = "#1f77b4", "Global" = "#ff7f0e")) +
  scale_x_continuous(breaks = seq(1990, 2020, 5)) +
  ylim(0, max(c(germany_data$Renewable_Energy_Share, 
                worldwide_avg$Renewable_Energy_Share), na.rm = TRUE) * 1.1)
```

## Additional Analysis

```{r}
# Calculate growth rates
germany_start <- germany_data$Renewable_Energy_Share[germany_data$Year == 1990]
germany_end <- germany_data$Renewable_Energy_Share[germany_data$Year == 2020]
germany_growth <- ((germany_end - germany_start) / germany_start) * 100

global_start <- worldwide_avg$Renewable_Energy_Share[worldwide_avg$Year == 1990]
global_end <- worldwide_avg$Renewable_Energy_Share[worldwide_avg$Year == 2020]
global_growth <- ((global_end - global_start) / global_start) * 100

# Create summary statistics table
summary_stats <- data.frame(
  Region = c("Germany", "Global"),
  Start_1990 = c(germany_start, global_start),
  End_2020 = c(germany_end, global_end),
  Absolute_Change = c(germany_end - germany_start, global_end - global_start),
  Percentage_Growth = c(germany_growth, global_growth)
)

print("Summary Statistics (1990-2020):")
print(summary_stats)

# Calculate correlation between Germany and global trends
correlation <- cor(germany_data$Renewable_Energy_Share, 
                   worldwide_avg$Renewable_Energy_Share, 
                   use = "complete.obs")
print(paste("Correlation between Germany and Global trends:", round(correlation, 3)))
```

# Interpretation

## Key Findings

The analysis reveals several important insights about renewable energy development:

1. **Germany's Leadership Role**: Germany demonstrates a clear pioneering position in renewable energy adoption, with consistently higher renewable energy shares compared to the global average throughout the entire period.

2. **Accelerated Growth**: Germany shows significant acceleration in renewable energy adoption, particularly from the early 2000s onwards, likely driven by policy initiatives such as the Renewable Energy Act (EEG).

3. **Global Progress**: While the global average shows more modest growth, there is still a positive trend, indicating worldwide recognition of the importance of renewable energy sources.

## Statistical Evidence

- Germany's renewable energy share increased substantially from 1990 to 2020
- The global average shows steady but slower growth over the same period
- The correlation analysis indicates the relationship between German and global trends
- Germany consistently outperforms the global average by a significant margin

## Conclusion

Germany has indeed taken a pioneering role in renewable energy development, significantly outpacing global trends. This leadership position is evident not only in absolute values but also in the rate of growth and sustained commitment to renewable energy expansion over three decades.

The data supports the hypothesis that Germany serves as a leader rather than a follower in the global renewable energy transition, potentially serving as a model for other nations seeking to increase their renewable energy adoption.
